---
title: "gtsummary in quarto"
format: 
  html:
    theme: minty
execute:
  echo: false
---

```{r}
#| message: false
library(tidyverse)
library(gtsummary)
```


```{r}
nlsy_cols <- c("glasses", "eyesight", "sleep_wkdy", "sleep_wknd", 
               "id", "nsibs", "samp", "race_eth", "sex", "region", 
               "income", "res_1980", "res_2002", "age_bir")
nlsy <- read_csv(here::here("data", "raw", "nlsy.csv"), 
                 na = c("-1", "-2", "-3", "-4", "-5", "-998"),
                 skip = 1, col_names = nlsy_cols) |> 
  mutate(region_cat = factor(region, labels = c("Northeast", "North Central", "South", "West")),
         sex_cat = factor(sex, labels = c("Male", "Female")),
         race_eth_cat = factor(race_eth, labels = c("Hispanic", "Black", "Non-Black, Non-Hispanic")),
         eyesight_cat = factor(eyesight, labels = c("Excellent", "Very good", "Good", "Fair", "Poor")),
         glasses_cat = factor(glasses, labels = c("No", "Yes"))) 
```

The NLSY79 Cohort is a longitudinal project that follows the lives of a sample of American youth born between 1957-64. The cohort originally included 12,686 respondents ages 14-22 when first interviewed in 1979; after two subsamples were dropped, 9,964 respondents remain in the eligible samples. Data are now available from Round 1 (1979 survey year) to Round 29 (2020 survey year).

# Customization of `tbl_summary()`

```{r}
#| label: tbl-one
#| tbl-cap: "Study Population Characteristics, by Sex"
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir))
```

@tbl-one shows descriptive statistics about the study population and @tbl-two shows the same data but with the variable names edited.


```{r}
#| label: tbl-two
#| tbl-cap: "Study Population Characteristics, by Sex"
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir),
  label = list(
    race_eth_cat ~ "Race/ethnicity",
    region_cat ~ "Region",
    eyesight_cat ~ "Eyesight",
    glasses ~ "Wears glasses",
    age_bir ~ "Age at first birth"
  ),
  missing_text = "Missing")
```

# Inline Text Practice
Create a list of overall stats you can refer to in a summary of the data

``` {r}
stats <- list(n = nrow(nlsy),
              females = sum(nlsy$sex == 2),
              males = sum(nlsy$sex == 1),
              mean_sleep_wknd = mean(nlsy$sleep_wknd, na.rm=TRUE),
              median_sleep_wknd = median(nlsy$sleep_wknd, na.rm=TRUE)
              )
```

There were `r stats$n` participants with an average sleep time of `r stats$mean_sleep_wknd` on the weekends. 


# Class Exercise 

Add a chunk with the table1 table from the slides.
Practice extracting the output using inline_text() as in the slides.
Edit or create a new table that has the median hours of sleep on weekends in it (in the overall sample).
From that table, pull out the that value and include in the text using inline_text().

```{r}
#| label: tbl-three
#| tbl-cap: "Study Population Characteristics (Median), by Sex"
tbl_summary(
  nlsy,
  include = sleep_wknd,
  label = list(sleep_wknd ~ "Hours slept on Weekends (Median)"),
  missing_text = "Missing",
	statistic = list(
		sleep_wknd = "{median}")
	)
```

@tbl-three shows that the median hours slept on weekends by the study participants was `r stats$median_sleep_wknd`
